Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-17 mu/article/article-review6-Anatoli.tex
Made Anatoli's changes.

Graphene, an allotrope of carbon with hexagonal 2D lattice structure, demonstrates 
properties such as fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08}, being a
perfect platform  for two-dimension (2D) electronic systems; however, most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that a narrow band gap  can be opened in graphene by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
Depending on the percentage of hydrogenation and spatial arrangements of
the hydrogen-carbon bonds, hydrogenated graphene demonstrates  different structural  
configurations and a tunable electron gap, as it has been proven in [1].
In this paper, we present two 50\% hydrogenated graphene noncentrosymmetric
structures, both presenting a discernible band gap. The first one, labelled as the Up structure (also known as graphone, see [2] and Refs therein), shown in
Fig. \ref{fig:up-struc}. Graphone has hydrogen atoms bonded to the carbon layer only on
the upper side of the structure, we consider here the magnetic isomer of graphone, with the so-called “chair” structure. In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/upstruc2}
    \\
    \includegraphics[width=\tama]{figures/upstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is also shown.
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/altstruc2}
    \\
    \includegraphics[width=\tama]{figures/altstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the 
      $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive rectangular unit cell is also displayed.

}
    \label{fig:alt-struc}
\end{figure}

@@

\hat K^{\mathrm{a}\mathrm{b}}=\frac{1}{2}\left( \hat v^\mathrm{a} \hat S^\mathrm{b} 

@@

%\end{document}

@@

{\mathbf E}(t)= {\mathbf E}(\omega)e^{-i\omega t} + {\mathbf E}^*(\omega)e^{i\omega t}

@@

In order to 
calculate the response of the system to ${\mathbf E}(t)$, one needs to
take into account the excited coherent superposition
of the spin-split conduction bands inherent to the 
noncentrosymmetric 
semiconductors considered in this work.
%Although, this splitting is small [15,16], is required to obtain the
%correct results\cite{nastosPRB07}.
% The conduction bands in the 
% noncentrosymmetric  
% semiconductors  
% are spin split by a small amount,[15,16] typically
%   smaller than the energy width of the laser pulse, and so the pulse
%   excites a coherent superposition of the two conduction bands. Even
%   for very long pulses with narrow energy widths, dephasing effects
%   lead to an energy width of the bands large enough that spin-split
%   states can become quasi-degenerate. 

@@

where we assumed that the conduction bands $c$ and $c'$
are quasi-degenerate  states, and we take $\epsilon\to 0$ at the end of the
calculation.  The spin-splitting of the valence ($v$) bands is very
small, and is neglected throughout this work.\cite{nastosPRB07}
The matrix elements of any operator ${\cal O}$ are given by
${\cal O}_{nm}({\mathbf k})=\langle n{\mathbf k}|\hat{\cal O}| m{\mathbf k}\rangle$,
where 
$H_{0}|n{\mathbf k}\rangle = \hbar \omega_{n}({\mathbf k})|n{\mathbf k}\rangle$ 
with $\hbar \omega_{n}({\mathbf k})$ being the energy of the
electronic band $n$ at point ${\mathbf k}$ in the irreducible Brillouin zone
(IBZ),  $|n{\mathbf k}\rangle$ is the Bloch state, and 
$\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf k})-\omega_{m}({\mathbf k})$.
Using 
$\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$
for the expectation value of an observable $\mathcal{O}$, 
where $\mathrm{Tr}$ denotes the trace, we obtain

@@

\mathcal{O} 
%=&
%\int \frac{d^{3}k}{8\pi^{3}}\sum_{c} \langle c {\mathbf k} 
%| \hat{\rho} \hat{\mathcal{O}} |
%c{\mathbf k} \rangle \nonumber \\
=& 

@@

where we used 
the closure
relationship $\sum_{n}|n{\mathbf k}\rangle \langle n{\mathbf k}| = 1$,
where $n=v,c$,
and the fact that $\rho_{vn}({\mathbf k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. 
Therefore, 
using  Eqs. \eqref{z.4} and \eqref{z.5},
the
rate of change of $\mathcal{O}$,
$\dot{\mathcal{O}} 
\mathrm{Tr} \left( \dot{\hat\rho} \hat{\cal O} \right)
$,
 is given by

@@

The prime symbol $'$ in the sum means that
$c$ and $c'$ are quasi-degenerate states, and the sum only covers these states.

Replacing  $\hat{\cal O} \rightarrow \hat{K}^{\mathrm{ab}}$, in 
the above expression, one can show that
\begin{equation}
\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
\label{eq:dotk}
\end{equation}
where repeated Cartesians are summed, and 
\begin{equation}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}} \sum'_{vcc'}
  \delta(\omega-\omega_{cv}({\mathbf k}) 
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 
\left(  
r^{\mathrm{c}}_{vc'}({\mathbf k})   
r^{\mathrm{d}}_{cv }({\mathbf k})  +
(c \leftrightarrow d)  
\right) 
\right]
\end{aligned}
\end{equation} 
is the pseudotensor that describes the rate of change of the  PSC process in
semiconductors. To derive what we presented above we used
$
K^{\mathrm{ab}}_{nm}({\mathbf -k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k}) 
$
which follows from time-reversal invariance. 
Since $\mu^{\mathrm{abcd}}(\omega)$ is real, we have that
$\mu^{\mathrm{abcd}}(\omega) =
\mu^{\mathrm{abdc}}(\omega)$. 
We point out that Eq.~\eqref{eq:mu} 
is –identical to Eq. (3) of Bhat et al.\cite{bhatPRL05},
derived using the semiconductor optical Bloch equations.
Using the closure relation,
\begin{equation}
K^{\mathrm{ab}}_{cc'}({\mathbf k}) = \frac{1}{2}
\sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}({\mathbf k})S^{\mathrm{b}}_{lc'}({\mathbf k})
+S^{\mathrm{b}}_{cl}({\mathbf k}) v^{\mathrm{a}}_{lc'}({\mathbf k})
\right)
,
\label{eq:velspimatelem}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $a$, % $\mathrm{a}$, 
at which the spin
moves in a polarized state along direction $b$. % $\mathrm{b}$. 
The carrier injection rate
$\dot n(\omega)$ is written as\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k})), 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\epsilon^{\mathrm{a}\mathrm{b}}(\omega)]=2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Fixing vel %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$,
we define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
defined with a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
= 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   R E S U L T S   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H & -0.61516 & -1.77416 &  0.73196 \\
H &  0.61518 &  0.35514 &  0.73175 \\
C & -0.61516 & -1.77264 & -0.49138 \\
C & -0.61516 & -0.35600 & -0.72316 \\
C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Up structure as the structural input for our calculations (the primitive hexagonal unit cell contains two C and one H atoms, see Fig. 1). ADD EXTRA C atom
}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H &  -0.61516 &  -1.42140 & \ 1.47237 \\
C &  -0.61516 &  -1.73300 & \ 0.39631 \\
C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
C &  -0.61516 &  -0.37396 &  -0.39632 \\
H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Alt structure, which coincides with its primitive cell. 
}
\label{tab:alt-unitcell}
\end{table}

@@

We calculated the self-consistent ground state and the Kohn-Sham states within 
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations,  corresponding to
the Up and Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting
in LDA energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 ${\mathbf k}$ points and 14452 ${\mathbf k}$ points in the IBZ; to
integrate $\mu^{\mathrm{abcd}}(\omega)$ and $\xi^{\mathrm{a}\mathrm{b}}(\omega)$ the
linearized analytic tetrahedron method (LATM) has been used.\cite{nastosPRB07} We neglect the
anomalous velocity term $\hbar(\boldsymbol{\sigma}\times\nabla V)/4m^2c^2$, where $V$ is the
crystal potential, in $\hat{\mathbf v}$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore,
$[\hat{\mathbf v},\hat{\mathbf S}]=0$, and Eq.~\eqref{z.1} reduces to $\hat K^{\mathrm{a}\mathrm{b}}=\hat
v^\mathrm{a} \hat S^\mathrm{b}=\hat S^\mathrm{b} \hat v^\mathrm{a}$. Finally, the prime in the sum of
Eq.~\eqref{eq:mu} is restricted to quasi-degenerated conduction bands $c$ and
$c'$ that are closer than 30 meV, which  is both typical laser-
pulse energy width and the thermal room-temperature energy level broadening.\cite{nastosPRB07}
% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% Results: Spin velocity %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% Res: mu & V^{ab} comparison %%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\includegraphics[width=\tama]{plots/2-vab-str-comp}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  vs. $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The largest velocity are at the low
  energy regions of the spectra for the Alt and Up structures, starting (going up?) at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to the 3D case of CdSe and GaAsThe, shown at the bottom panel . } 
\label{fig:vab-str-comp}
\end{figure}

>>>>

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimension (2D)
electronic systems; however, most electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} boron-
nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} noble-
metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in [1]. In this paper, we present two 50\% hydrogenated graphene
noncentrosymmetric structures, both presenting a discernible band gap. The
first one, labelled as the Up structure (also known as graphone, see [2] and
Refs therein), shown in Fig. \ref{fig:up-struc}. Graphone has hydrogen atoms
bonded to the carbon layer only on the upper side of the structure, we consider
here the magnetic isomer of graphone, with the so-called “chair” structure. In
contrast,  the Alt structure, shown in Fig. \ref{fig:alt-struc}, has hydrogen
alternating on the upper and bottom sides of the carbon
sheet.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/fig1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is also shown.
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/fig2}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the $xyz$ directions. The dark (light) spheres are
      the C (H) atoms. The primitive rectangular unit cell is also displayed.}
    \label{fig:alt-struc}
\end{figure}

@@

\hat K^{\mathrm{a}\mathrm{b}} = 
\frac{1}{2}\left( \hat v^\mathrm{a} \hat S^\mathrm{b} 

@@

@@

{\mathbf E}(t) = 
{\mathbf E}(\omega)e^{-i\omega t} + {\mathbf E}^*(\omega)e^{i\omega t}

@@

In order to calculate the response of the system to ${\mathbf E}(t)$, one needs
to take into account the excited coherent superposition of the spin-split
conduction bands inherent to the noncentrosymmetric semiconductors considered
in this work.
%Although, this splitting is small [15,16], is required to obtain the correct
%results\cite{nastosPRB07}. The conduction bands in the noncentrosymmetric
%semiconductors are spin split by a small amount,[15,16] typically smaller than
%the energy width of the laser pulse, and so the pulse excites a coherent
%superposition of the two conduction bands. Even for very long pulses with
%narrow energy widths, dephasing effects lead to an energy width of the bands
%large enough that spin-split states can become quasi-degenerate.

@@

where we assumed that the conduction bands $c$ and $c'$ are quasi-degenerate
states, and we take $\epsilon\to 0$ at the end of the calculation.  The spin-
splitting of the valence ($v$) bands is very small, and is neglected throughout
this work.\cite{nastosPRB07} The matrix elements of any operator ${\cal O}$ are
given by ${\cal O}_{nm}({\mathbf k})=\langle n{\mathbf k}|\hat{\cal O}|
m{\mathbf k}\rangle$, where $H_{0}|n{\mathbf k}\rangle = \hbar
\omega_{n}({\mathbf k})|n{\mathbf k}\rangle$ with $\hbar \omega_{n}({\mathbf
k})$ being the energy of the electronic band $n$ at point ${\mathbf k}$ in the
irreducible Brillouin zone (IBZ),  $|n{\mathbf k}\rangle$ is the Bloch state,
and $\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf k})-\omega_{m}({\mathbf k})$.
Using $\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$ for the
expectation value of an observable $\mathcal{O}$, where $\mathrm{Tr}$ denotes
the trace, we obtain

@@

\mathcal{O} = & 

@@

where we used the closure relationship $\sum_{n}|n{\mathbf k}\rangle \langle
n{\mathbf k}| = 1$, where $n=v,c$, and the fact that $\rho_{vn}({\mathbf
k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. Therefore, using  Eqs. \eqref{z.4}
and \eqref{z.5}, the rate of change of $\mathcal{O}$, $\dot{\mathcal{O}} =
\mathrm{Tr} \left( \dot{\hat\rho} \hat{\cal O} \right)$, is given by

@@

The prime symbol $'$ in the sum means that $c$ and $c'$ are quasi-degenerate
states, and the sum only covers these states. Replacing  $\hat{\cal O}
\rightarrow \hat{K}^{\mathrm{ab}}$, in the above expression, one can show that
\begin{equation}
\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
\label{eq:dotk}
\end{equation}
where repeated Cartesians are summed, and 
\begin{equation}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}} \sum'_{vcc'}
\delta(\omega-\omega_{cv}({\mathbf k}) 
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 
\left(  
r^{\mathrm{c}}_{vc'}({\mathbf k})   
r^{\mathrm{d}}_{cv }({\mathbf k})  +
(c \leftrightarrow d)  
\right) 
\right]
\end{aligned}
\end{equation} 
is the pseudotensor that describes the rate of change of the  PSC process in
semiconductors. To derive what we presented above we used $
K^{\mathrm{ab}}_{nm}({\mathbf -k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$ which
follows from time-reversal invariance. Since $\mu^{\mathrm{abcd}}(\omega)$ is
real, we have that $\mu^{\mathrm{abcd}}(\omega) =
\mu^{\mathrm{abdc}} (\omega)$. We point out that Eq.~\eqref{eq:mu} is
–identical to Eq. (3) of Bhat et al.\cite{bhatPRL05}, derived using the
semiconductor optical Bloch equations. Using the closure relation,
\begin{equation}
K^{\mathrm{ab}}_{cc'}({\mathbf k}) = \frac{1}{2}
\sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}({\mathbf k})S^{\mathrm{b}}_{lc'}({\mathbf k})
+S^{\mathrm{b}}_{cl}({\mathbf k}) v^{\mathrm{a}}_{lc'}({\mathbf k})
\right)
\label{eq:velspimatelem}
\end{equation}

We define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $a$, at which the spin moves in a
polarized state along direction $b$. The carrier injection rate $\dot
n(\omega)$ is written as\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k})), 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical response tensor by
$\mathrm{Im} [\epsilon^{\mathrm{a}\mathrm{b}}(\omega)] =
2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

@@

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system's
coordinates defined with a polar angle
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and an azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H & -0.61516 & -1.77416 &  0.73196 \\
H &  0.61518 &  0.35514 &  0.73175 \\
C & -0.61516 & -1.77264 & -0.49138 \\
C & -0.61516 & -0.35600 & -0.72316 \\
C &  0.61516 &  0.35763 & -0.49087 \\
C &  0.61518 &  1.77416 & -0.73196 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Up structure as the structural
input for our calculations (the primitive hexagonal unit cell contains two C
and one H atoms, see Fig. 1). ADD EXTRA C atom }
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H &  -0.61516 &  -1.42140 & \ 1.47237 \\
C &  -0.61516 &  -1.73300 & \ 0.39631 \\
C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
C &  -0.61516 &  -0.37396 &  -0.39632 \\
H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Alt structure, which coincides
with its primitive cell. }
\label{tab:alt-unitcell}
\end{table}

@@

We calculated the self-consistent ground state and the Kohn-Sham states within
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used Hartwigsen-
Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented in Eq.
\eqref{eq:mu}.
The convergence parameters for the calculations,  corresponding to the Up and
Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting in LDA
energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 ${\mathbf k}$ points and 14452 ${\mathbf k}$ points in the IBZ; to
integrate $\mu^{\mathrm{abcd}}(\omega)$ and
$\xi^{\mathrm{a}\mathrm{b}}(\omega)$ the linearized analytic tetrahedron method
(LATM) has been used.\cite{nastosPRB07} We neglect the anomalous velocity term
$\hbar(\boldsymbol{\sigma}\times\nabla V)/4m^2c^2$, where $V$ is the crystal
potential, in $\hat{\mathbf v}$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore, $[\hat{\mathbf
v},\hat{\mathbf S}]=0$, and Eq.~\eqref{z.1} reduces to $\hat
K^{\mathrm{a}\mathrm{b}}=\hat v^\mathrm{a} \hat S^\mathrm{b}=\hat S^\mathrm{b}
\hat v^\mathrm{a}$. Finally, the prime in the sum of Eq.~\eqref{eq:mu} is
restricted to quasi-degenerated conduction bands $c$ and $c'$ that are closer
than 30 meV, which  is both typical laser- pulse energy width and the thermal
room-temperature energy level broadening.\cite{nastosPRB07}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  vs. $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, starting (going up?) at the energy gap of each structure. In
  the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAsThe, shown at the bottom panel . }
\label{fig:vab-str-comp}
\end{figure}
